---
title: "Landscape PCA"
author: "Fonti Kar"
date: "8/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, include = T)
rm(list=ls())
sessionInfo()
#R version 3.6.1 (2019-04-26)
#Platform: x86_64-apple-darwin15.6.0 (64-bit)
#Running under: macOS High Sierra 10.13.6

pacman::p_load(tidyverse, magrittr, dplyr, ggplot2, GGally, MASS, psych, Matrix)
```

## Read in data

```{r data read}
landdat <- read.csv("data/landscape_dat.csv", stringsAsFactors = F)
str(landdat)
names(landdat)
```

## Tidying up some variable names

```{r}
landdat %<>% rename(site = Site.name,
                    site_type = Type,
                    Num_spiders = X..spiders,
                    Percent_vegetation = X.vegetation,
                    Percent_grass = X.grass,
                    Percent_hardsurfance = X.hard.surface,
                    Percent_water = X.water,
                    Ave_week_income = Ave..weekly.income)

head(landdat)
```

## Correcting Hoebeck's missing data

```{r}
landdat %>% filter(site == "Hoebeck")

landdat %<>% mutate(living.area.per.person = ifelse(site == "Hoebeck", 0, living.area.per.person),
                   Number.of.housing.complexes = ifelse(site == "Hoebeck", 0, Number.of.housing.complexes),
                   Number.of.dwellings = ifelse(site == "Hoebeck", 0, Number.of.dwellings),
                   Ave_week_income = ifelse(site == "Hoebeck", 0, Ave_week_income))
```

## Correlation among variables?

```{r}
landvars <- landdat %>% dplyr::select(Percent_vegetation:Ave_week_income)

ggpairs(landvars, palette = "RdBu", label = TRUE, hjust = 0.75, size = 3, layout.exp = 1)
```

## Standardise and transform variables

```{r}
scaled_landdat <- landdat %>% mutate_at(names(landvars), scale)

sc_landvars <- scaled_landdat %>% dplyr::select(Percent_vegetation:Ave_week_income)
```

## Correlations of scaled variables

http://www.sthda.com/english/wiki/print.php?id=206

```{r}
land.cor <- cor(sc_landvars)

# Bartlett's test
cortest.bartlett(sc_landvars, n = 9)

#Steiger's test
cortest.mat(cor(sc_landvars), n1 = 9)
 
# KMO(Kaiser-Meyer-Olkin)</span>
KMO(cor(sc_landvars))
 
# Determinant
det(cor(sc_landvars))
```

#Do the PCA on scaled variables
Useful resources: 

https://www.analyticsvidhya.com/blog/2016/03/practical-guide-principal-component-analysis-python/

http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/

Good for checks:
https://www.r-bloggers.com/naive-principal-component-analysis-using-r/

##Using the psych package

```{r}
pc1 = principal(sc_landvars, nfactors = 
length(sc_landvars), rotate="varimax", scores = T)

pc1

str(pc1)

#Biplot
biplot(pc1)

#screen plot
plot(pc1$values, type ="b") 

#health check
pc1$residual %>% plot
pc1$fit
pc1$communality 

#Put these PC scores back into original dataframe but first check it matches number of data points to number of rows in data
length(prin1.0$x[,1])
nrow(scaled_landdat)

scaled_landdat$landPC1 <- pc1$scores[,1] #Strongly loaded on Percent_hardsurfance, Total.population.of.suburb, Population.density.people.km2, Num dwellings

#write.csv(scaled_landdat, row.names = F, "output/sc_landscapdatPC.csv")

arrange(scaled_landdat, site_type) %>% dplyr::select(site, site_type, landPC1)
```

## Take out the money and human related variables because they are not matching up!

```{r}
head(sc_landvars)

sc_land_subset <- sc_landvars %>% dplyr::select(Percent_vegetation:site.distance.to.water)
```

#Pre-diagnoses
```{r}
sc_land_subset.cor <- psych::corr.test(sc_land_subset)

# Bartlett's test
cortest.bartlett(sc_land_subset.cor, n = nrow(sc_land_subset.cor))

#Steiger's test
cortest.mat(cor(sc_land_subset), n1 = length(cortest.mat))
 
# KMO(Kaiser-Meyer-Olkin)</span>
KMO(cor(sc_land_subset))
 
# Determinant
det(cor(sc_land_subset))
```

#Do the PCA on scaled variables of the subset (no human variables)

No this is performing worst, not ideal at all.
I will suggest using a score I think! Or the Categories she already has

```{r}
pc2 = psych::principal(sc_land_subset, nfactors = 
length(sc_land_subset), rotate="varimax", scores = T)

pc2 = pca(sc_land_subset, nfactors = 
length(sc_land_subset), rotate="varimax", scores = T)

pc2

str(pc2)

#Biplot
biplot(pc2)

#screen plot
plot(pc2$values, type ="b") 

#health check
pc2$residual %>% plot
pc2$fit
pc2$communality 

#Put these PC scores back into original dataframe but first check it matches number of data points to number of rows in data
scaled_landdat$landsub_pc1 <- pc2$scores[,1] #Strongly loaded by Percent_water and site.distance.to.forrest 
arrange(scaled_landdat, site_type) %>% dplyr::select(site, site_type, landsub_pc1) #The most rural spots are not getting the highest values in PC scores
```

#Trying another PCA function
```{r}
prcomp(landvars) %>% summary()
prcomp(landvars) %>% plot()
prcomp(landvars) %>% biplot()

princomp(landvars) %>% summary()
princomp(landvars) %>% plot()
princomp(landvars) %>% biplot()
princomp(landvars)$loadings

names(scaled_landdat)
```

#Merge into data_core data
```{r}
data_core <- read.csv("data/processed/data_core.csv", stringsAsFactors = F)

str(data_core)
names(data_core)
head(data_core)

str(scaled_landdat)

#Match site names
data_core$site %>% unique() %>% sort() == scaled_landdat$site %>% unique() %>% sort()

#Fix typos
scaled_landdat %<>% mutate(site = ifelse(site == "Barmbeck Sud", "Barmbek Sud", scaled_landdat$site))
scaled_landdat %<>% mutate(site = ifelse(site == "Botanical Gardens", "Botanical gardens", scaled_landdat$site))
scaled_landdat %<>% mutate(site = ifelse(site == "Hesse Park", "Hasse Park", scaled_landdat$site))
scaled_landdat %<>% mutate(site = ifelse(site == "Staarlandsrasse", "Staarlandstrasse", scaled_landdat$site))
scaled_landdat %<>% mutate(site = ifelse(site == "Staats Forest", "Staats forest", scaled_landdat$site))

#Merge in PCA and land data
data_all <- left_join(data_core, scaled_landdat %>% dplyr::select(site, landPC1, landsub_pc1))

write.csv(data_all, row.names = F, "data/processed/data_all.csv")
```

#Merge with lab_long_data
```{r}
lab_data_long <-read.csv("data/processed/lab_data_long.csv", stringsAsFactors = F)

str(lab_data_long)
names(lab_data_long)
head(lab_data_long)

str(scaled_landdat)

#Match site names
lab_data_long$site %>% unique() %>% sort() == scaled_landdat$site %>% unique() %>% sort()

#Merge in PCA and land data
lab_data_long_all <- left_join(lab_data_long, scaled_landdat %>% dplyr::select(site, landPC1, landsub_pc1))

str(lab_data_long_all)

write.csv(lab_data_long_all, row.names = F, "data/processed/lab_data_long_all.csv")
```


#Explore relationship with PC scores and urban categories
```{r}
#Morpho
ggplot(data_all, aes(x = landPC1, y = z_ln_weight, label = site)) + 
  geom_point(colour = "white") + 
  geom_text() + 
  stat_smooth(method = "lm")

ggplot(data_all, aes(x = landsub_pc1, y = z_ln_weight, label = site)) + 
  geom_point(colour = "white") + 
  geom_text() + 
  stat_smooth(method = "lm")

#Behaviour
ggplot(data_all, aes(x = landPC1, y = z_ln_Tactive, label = site_type)) + 
  geom_point(colour = "white") + 
  geom_text() + 
  stat_smooth(method = "lm")

ggplot(data_all, aes(x = landPC1, y = z_poke_score, label = site_type)) + 
  geom_point(colour = "white") + 
  geom_text() + 
  stat_smooth(method = "lm")

ggplot(data_all, aes(x = landPC1, y = z_ln_tot_prey_attack, label = site_type)) + 
  geom_point(colour = "white") + 
  geom_text() + 
  stat_smooth(method = "lm")
```

