---
title: "Main analyses"
author: "Fonti Kar"
date: "9/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, include = T)
rm(list=ls())
sessionInfo()
# R version 3.6.1 (2019-07-05)
# Platform: x86_64-apple-darwin15.6.0 (64-bit)
# Running under: macOS High Sierra 10.13.6

pacman::p_load(tidyverse, magrittr, dplyr, ggplot2, GGally, patchwork, MCMCglmm, brms, janitor)
```

## Read in combined data

```{r data}
data <- read.csv("data/processed/data_for_analysis.csv", stringsAsFactors = F)
data_long <- read.csv("data/processed/data_long_analysis.csv", stringsAsFactors = F)
data_all <- read.csv("data/processed/data_allvars.csv") #for descirptive stats calculations

str(data)
head(data)

str(data_long)
head(data_long)

str(data_all)
head(data_all)
```

## Fix NA sex
```{r}
# data %>% filter(is.na(sex))
# data %>% filter(sex == "I")
# data %>% filter(sex == "F")
# data %>% filter(sex == "M")
data %<>% mutate(sex = ifelse(is.na(sex), "I", data$sex))
data_long %<>% mutate(sex = ifelse(is.na(sex), "I", data$sex))
```

## 1) Descriptive stats: What are the overall trait means and variances? 

```{r}
data_all %>% select(poke_score, tot_prey_attack, Tactive) %>% summary()

data_all %>% select(poke_score, tot_prey_attack, Tactive) %>% 
  summarise(mu_poke = mean(poke_score, na.rm = T),
            mu_tot_prey_attack = mean(tot_prey_attack, na.rm = T),
            mu_Tactive = mean(Tactive, na.rm = T))

data_all %>% select(poke_score, tot_prey_attack, Tactive) %>% 
  summarise(sd_poke = sd(poke_score, na.rm = T),
            sd_tot_prey_attack = sd(tot_prey_attack, na.rm = T),
            sd_Tactive = sd(Tactive, na.rm = T))
```

### How many measures for each trait for each individual
```{r}
data_long %>% group_by(spider_id) %>% summarise(n = length(spider_id)) %>% as.data.frame() # 6 is good
```

##  2)  Are traits overall repeatable?
Need to use data_long for these analyses not data! Can only do with traits with more than 2 data points per ID. Can't do for percent movement. 

Yees traits are overall moderately repeatable irrespective of urban site type (i.e spiders exhibit 'personalities' and individuals consistenly differ in how their behave (poke and Tactive only))

```{r}
mod.1 <- brm(z_poke_score ~ sex + site_type + z_days_noweb + (1|spider_id),
    family = gaussian(),
    data = data_long, 
    chains = 4, cores = 2, iter = 6500, thin = 4, warmup = 1500, 
    control = list(adapt_delta = 0.99))

plot(mod.1)
summary(mod.1)

mod.2 <- brm(z_Tactive ~ sex + site_type + z_days_noweb + (1|spider_id),
    family = gaussian(),
    data = data_long, 
    chains = 4, cores = 2, iter = 5500, thin = 4, warmup = 1500, 
    control = list(adapt_delta = 0.98))

summary(mod.2)
```

### MCMCglmm model 
```{r}
# Set up priors for model and configure model
prior <- list(G = list(G1 =list(V = 1, nu = 0.0002)),
              R = list(V = 1, nu = 0.002))

nitt = 553000
burnin = 3000
thin = 200

mod.1 <- MCMCglmm(z_poke_score ~ sex + site_type + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = T)
         
saveRDS(mod.1, "output/rds/mod.1")
mod.1 <- readRDS("output/rds/mod.1")
plot(mod.1)
summary(mod.1)

#Repeatabilty of poke score is calculated as:
#Point estimate
0.2286/(0.2286+0.7592) #Pretty moderately repeatable
poke_score_rep <- mod.1$VCV[,"spider_id"]/(mod.1$VCV[,"spider_id"] + mod.1$VCV[,"units"]) 
posterior_summary(poke_score_rep) #Point estimate and credible intervals

mod.2 <- MCMCglmm(z_ln_Tactive_lab ~ sex + site_type + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = T)

saveRDS(mod.2, "output/rds/mod.2")
mod.2 <- readRDS("output/rds/mod.2")
plot(mod.2)
summary(mod.2)

#Repeatability of Tactive is calculated as: 
Tactive_score_rep <- mod.2$VCV[,"spider_id"]/(mod.2$VCV[,"spider_id"] + mod.2$VCV[,"units"]) 
posterior_summary(Tactive_score_rep) #Point estimate and credible intervals

```

## 3) Does repeatability vary with site categories? 
Subset data and run models
Credible intervals all overlap, but maybe sub urban is a bit more variable
```{r}
#How much data for each site category?
data_long %>% filter(site_type == "Rural") %>% nrow()
data_long %>% filter(site_type == "Sub urban") %>% nrow()
data_long %>% filter(site_type == "Urban") %>% nrow()

rural_data_long <- data_long %>% filter(site_type == "Rural")
sub_data_long <- data_long %>% filter(site_type == "Sub urban")
urb_data_long <- data_long %>% filter(site_type == "Urban") 

#Poke score
mod.1_rural <- MCMCglmm(z_poke_score ~ sex  + z_days_noweb + z_ln_weight,
                        random = ~spider_id,
                        prior = prior,
                        nitt = nitt,
                        burnin = burnin,
                        thin = thin,
                        data = rural_data_long, 
                        verbose = T)

saveRDS(mod.1_rural, "output/rds/mod.1_rural")
mod.1_rural <- readRDS("output/rds/mod.1_rural")
plot(mod.1_rural)
summary(mod.1_rural)

rural_poke_score_rep <- mod.1_rural$VCV[,"spider_id"]/(mod.1_rural$VCV[,"spider_id"] + mod.1_rural$VCV[,"units"]) 
posterior_summary(rural_poke_score_rep) #Point estimate and credible intervals

mod.1_sub <- MCMCglmm(z_poke_score ~ sex  + z_days_noweb + z_ln_weight,
                        random = ~spider_id,
                        prior = prior,
                        nitt = nitt,
                        burnin = burnin,
                        thin = thin,
                        data = sub_data_long, 
                        verbose = F)

saveRDS(mod.1_sub, "output/rds/mod.1_sub")
mod.1_sub <- readRDS("output/rds/mod.1_sub")
plot(mod.1_sub)
summary(mod.1_sub)

sub_poke_score_rep <- mod.1_sub$VCV[,"spider_id"]/(mod.1_sub$VCV[,"spider_id"] + mod.1_sub$VCV[,"units"]) 
posterior_summary(sub_poke_score_rep) #Point estimate and credible intervals

mod.1_urb <- MCMCglmm(z_poke_score ~ sex  + z_days_noweb + z_ln_weight,
                        random = ~spider_id,
                        prior = prior,
                        nitt = nitt,
                        burnin = burnin,
                        thin = thin,
                        data = urb_data_long, 
                        verbose = T)

saveRDS(mod.1_urb, "output/rds/mod.1_urb")
mod.1_urb <- readRDS("output/rds/mod.1_urb")
plot(mod.1_urb)
summary(mod.1_urb)

urb_poke_score_rep <- mod.1_urb$VCV[,"spider_id"]/(mod.1_urb$VCV[,"spider_id"] + mod.1_urb$VCV[,"units"]) 
posterior_summary(urb_poke_score_rep) #Point estimate and credible intervals

#Tactive
mod.2_rural <- MCMCglmm(z_ln_Tactive_lab ~ sex + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = rural_data_long, 
                  verbose = F)

saveRDS(mod.2_rural, "output/rds/mod.2_rural")
mod.2_rural <- readRDS("output/rds/mod.2_rural")
plot(mod.2_rural)
summary(mod.2_rural)

rural_Tactive_score_rep <- mod.2_rural$VCV[,"spider_id"]/(mod.2_rural$VCV[,"spider_id"] + mod.2_rural$VCV[,"units"]) posterior_summary(rural_Tactive_score_rep) #Point estimate and credible intervals

mod.2_sub <- MCMCglmm(z_ln_Tactive_lab ~ sex + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = sub_data_long, 
                  verbose = F)

saveRDS(mod.2_sub, "output/rds/mod.2_sub")
mod.2_sub <- readRDS("output/rds/mod.2_sub")
plot(mod.2_sub)
summary(mod.2_sub)

sub_Tactive_score_rep <- mod.2_sub$VCV[,"spider_id"]/(mod.2_sub$VCV[,"spider_id"] + mod.2_sub$VCV[,"units"]) 
posterior_summary(sub_Tactive_score_rep) #Point estimate and credible intervals

mod.2_urb <- MCMCglmm(z_ln_Tactive_lab ~ sex + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = urb_data_long, 
                  verbose = F)

saveRDS(mod.2_urb, "output/rds/mod.2_urb")
mod.2_urb <- readRDS("output/rds/mod.2_urb")
plot(mod.2_urb)
summary(mod.2_urb)

urb_Tactive_score_rep <- mod.2_urb$VCV[,"spider_id"]/(mod.2_urb$VCV[,"spider_id"] + mod.2_urb$VCV[,"units"]) 
posterior_summary(urb_Tactive_score_rep) #Point estimate and credible intervals
```


## 4) Are there differences in means trait values between site categories? 
Intercept = Rural
Releveled to get Intercept = Urban to get Suburban vs Urban constrast
No there are no differences in poke scores among site types

```{r}
summary(mod.1) #No diff with Rural vs Suburban, Rural vs Urban in poke score
summary(mod.2) #No diff with Rural vs Suburban, Rural vs Urban in Tactive

data_long %<>% mutate(site_type = as.factor(site_type))

#Relevel model with Urban as intercept
mod.1.5 <- MCMCglmm(z_poke_score ~ sex + relevel(site_type, ref = "Urban") + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = F)

saveRDS(mod.1.5, "output/rds/mod.1.5")
mod.1.5 <- readRDS("output/rds/mod.1.5")

plot(mod.1.5)
summary(mod.1.5)

mod.2.5 <- MCMCglmm(z_ln_Tactive_lab ~ sex + relevel(site_type, ref = "Urban") + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = F)

saveRDS(mod.2.5, "output/rds/mod.2.5")
mod.2.5 <- readRDS("output/rds/mod.2.5")

plot(mod.2.5)
summary(mod.2.5)
```

## 5) Does trait means change along urban index 
No signficant relationship with poke score along landscape index
Time active decreases along as landscape index increases (i.e. Total population increases) #moderately significant

```{r}
mod.3 <- MCMCglmm(z_ln_Tactive_lab ~ sex + landPC1 + z_days_noweb  + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = F)

saveRDS(mod.3, "output/rds/mod.3")
mod.3 <- readRDS("output/rds/mod.3")

plot(mod.3)
summary(mod.3)

mod.4 <- MCMCglmm(z_poke_score ~ sex + landPC1 + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = F)

saveRDS(mod.4, "output/rds/mod.4")
mod.4 <- readRDS("output/rds/mod.4")

plot(mod.4)
summary(mod.4)
```

## 6) Does trait means change along microhabitat index 
Neither traits varied with microhabitat index, coefficient is positive for poke score and negative for Tactive

```{r}
mod.5 <- MCMCglmm(z_ln_Tactive_lab ~ sex + mh_pc1 + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = F)

saveRDS(mod.5, "output/rds/mod.5")
plot(mod.5)
summary(mod.5)

mod.6 <- MCMCglmm(z_poke_score ~ sex + mh_pc1 + z_days_noweb + z_ln_weight,
                  random = ~spider_id,
                  prior = prior,
                  nitt = nitt,
                  burnin = burnin,
                  thin = thin,
                  data = data_long, 
                  verbose = F)

saveRDS(mod.6, "output/rds/mod.6")

plot(mod.6)
summary(mod.6)


```

## 7) Difference with field and lab
```{r}
names(data)
data %>% filter(in_situ == 0) %>% nrow()
data %>% filter(in_situ == 1) %>% nrow()

#Poke
complete_poke <- data[complete.cases(data$z_poke_score),]
nrow(complete_poke)

mod.7 <- glm(z_poke_score ~ sex + in_situ + z_ln_weight, data = complete_poke)
summary(mod.7)

#Tactive
complete_Tactive <- data[complete.cases(data$z_ln_Tactive),]
nrow(complete_Tactive)
summary(complete_Tactive)

mod.8 <- glm(z_ln_Tactive ~ sex + z_ln_weight, data = complete_Tactive)
summary(mod.8)
```

## 8) Differences in traits with site types that were only measured once
```{r}
#Prey attacked
complete_prey <- data[complete.cases(data$z_ln_tot_prey_attack),]
nrow(complete_prey)
summary(complete_prey)

mod.9 <- glm(z_ln_tot_prey_attack ~ sex + site_type + z_ln_weight, data = complete_prey)
summary(mod.9)
```

## 9) Do these traits on average change with microhab? Irrespective of ID variation essentially

```{r}
#Prey attacked
mod.10 <- glm(z_ln_tot_prey_attack ~ sex + landPC1 + z_ln_weight, data = complete_prey)
summary(mod.10)

#Tactive
mod.11 <- glm(z_ln_Tactive ~ sex + landPC1 + z_ln_weight, data = complete_Tactive)
summary(mod.11)

#Poke
mod.12 <- glm(z_poke_score ~ sex + landPC1 + z_ln_weight, data = complete_poke)
summary(mod.12)
```

## 10) Do these traits on average change with landscape? Irrespective of ID variation essentially

```{r}
#Prey attacked
mod.13 <- glm(z_ln_tot_prey_attack ~ sex + mh_pc1 + z_ln_weight, data = complete_prey)
summary(mod.13)

#Tactive
mod.14 <- glm(z_ln_Tactive ~ sex + mh_pc1 + z_ln_weight, data = complete_Tactive)
summary(mod.14)

#Poke
mod.15 <- glm(z_poke_score ~ sex + mh_pc1 + z_ln_weight, data = complete_poke)
summary(mod.15)
```

## 11) Correlations in traits at the population level among urban categories
```{r}
names(data)

data %>% filter(site_type == "Rural") %>% nrow()
data %>% filter(site_type == "Sub urban") %>% nrow()
data %>% filter(site_type == "Urban") %>% nrow()

data %>% select(c(z_poke_score, z_ln_tot_prey_attack, z_ln_Tactive))

mod.16 <- brm(mvbind(z_poke_score, z_ln_tot_prey_attack, z_ln_Tactive) ~ sex + z_ln_weight + (1|s|site_type), 
                   data = data,
              control = list(adapt_delta = 0.99),
              chains = 4, cores = 2, iter = 6500, thin = 4, warmup = 1500)

saveRDS(mod.16, "output/rds/mod.16")

summary(mod.16)
posterior_summary(mod.16)
```

